
---
title: "Cyclistic Bike Share Data Analysis"
author: "Ebrahim Isskandar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(googledrive)
```

## Introduction

This analysis explores bike ride data across four quarters. I aim to compare ride behaviors between different user types, understand usage patterns, and derive insights.

## Load and Clean Data

```{r load-data}
file_ids <- c(
  "1wDWoOtOS-7H9KjwpkczVOrMiJf8kAvBA",
  "17Ig2DZWLw5hVa3xIbOdVfYUhPZ1I9K1A",
  "11K3HQSRQMpeNc_maIDFU3vUUPxOkd5qy",
  "1Rqi4Eers94NxkAxglexNTGslaEmDY43b"
)

drive_auth()

for(i in seq_along(file_ids)){
  drive_download(as_id(file_ids[i]), path = paste0("q", i, ".csv"), overwrite = TRUE)
}

q1 <- read_csv("q1.csv")
q2 <- read_csv("q2.csv")
q3 <- read_csv("q3.csv")
q4 <- read_csv("q4.csv")

colnames(q2) <- colnames(q1)
colnames(q3) <- colnames(q1)
colnames(q4) <- colnames(q1)

all_data <- bind_rows(q1, q2, q3, q4)
```

```{r clean-data}
all_data <- all_data %>%
  mutate(
    start_time = ymd_hms(start_time),
    end_time = ymd_hms(end_time),
    ride_length = as.numeric(difftime(end_time, start_time, units = "mins"))
  ) %>%
  filter(ride_length > 0)
```

## Summary Statistics

```{r summary}
format(summary(all_data$ride_length), scientific = FALSE)
```

## Ride Count by User Type

```{r ride-count}
all_data %>%
  count(usertype) %>%
  ggplot(aes(x = usertype, y = n, fill = usertype)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Rides by User Type", x = "User Type", y = "Number of Rides") +
  theme_minimal()
```

## Average Ride Length by User Type

```{r avg-ride}
all_data %>%
  group_by(usertype) %>%
  summarise(avg_ride_length = mean(ride_length)) %>%
  ggplot(aes(x = usertype, y = avg_ride_length, fill = usertype)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Ride Length by User Type", x = "User Type", y = "Ride Duration (minutes)") +
  theme_minimal()
```

## Weekly Trends

```{r weekday-trend}
all_data <- all_data %>%
  mutate(day_of_week = wday(start_time, label = TRUE))

all_data %>%
  group_by(usertype, day_of_week) %>%
  summarise(trip_count = n(), .groups = "drop") %>%
  ggplot(aes(x = day_of_week, y = trip_count, fill = usertype)) +
  geom_col(position = "dodge") +
  labs(title = "Number of Rides per Day of Week by User Type", x = "Day", y = "Number of Rides") +
  theme_minimal()
```

## Ride Length Distribution

```{r boxplot}
all_data %>%
  ggplot(aes(x = usertype, y = ride_length, fill = usertype)) +
  geom_boxplot() +
  coord_cartesian(ylim = c(0, 100)) +
  labs(title = "Ride Length Comparison by User Type", x = "User Type", y = "Ride Duration (minutes)") +
  theme_minimal()
```

## Conclusion

This analysis highlights different riding patterns based on user type. Subscribers tend to ride shorter but more frequently. Understanding these behaviors can help optimize operations and marketing strategies.
